---
title: "Chapter 3"
author: "Carilli"
date: '`r format(lubridate::today(), "%B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    code_folding: hide
    toc: false
    toc_float: false
    theme: cerulean
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA,
                      options(digits = 3, scipen = 999))
library(tidyverse)
    
```

# Exercises {.tabset}

## 1 {.tabset}

```{r 1-data, message=FALSE}
pres_vote <-
  here::here("data", "PresVote.csv") %>% 
  read_csv()
```

### a

```{r 1a, message=FALSE}
pres_vote %>%
  ggplot(aes(x = rdi4, y = vote, label = year)) +
  geom_point() +
  labs(x = "Percent Change in Income",
       y = "Incumbent party's vote percent",
       caption = "Figure 3.1: Relationship between Income Growth and Vote fore the Incumbent President's Party.") +
  geom_smooth(method = "lm", se = FALSE) +
  theme(axis.line = element_line(),
        panel.background = element_blank()) +
  ggrepel::geom_text_repel()
```

### b

```{r 1b}
pres_vote %$% 
  lm(vote ~ rdi4) %>%
  jtools::summ()
```

### c

```{r 1c}
pres_vote %>% 
  lm(vote ~ rdi4, .) %>% 
  broom::augment() %>% 
  mutate(year = pres_vote$year) %>% 
  filter(year %in% c(1996, 1972)) %>% 
  select(year, .fitted)
  
```

### d

```{r 1d}
pres_vote %>% 
  lm(vote ~ rdi4, .) %>% 
  broom::augment() %>% 
  mutate(year = pres_vote$year) %>% 
  filter(year %in% c(1996, 1972)) %>% 
  select(year, .resid)
```

### e

```{r 1e}
pres_vote %>% 
  filter(reelection == 1) %>% 
  lm(vote ~ rdi4, .) %>% 
  jtools::summ()

```

### f

```{r 1f}
pres_vote %>% 
  filter(reelection == 0) %>% 
  lm(vote ~ rdi4, .) %>% 
  jtools::summ()
```



## 2 {.tabset}

### Program Monte Carlo OLS Simulation {.tabset}

The population regression function is $Salary = 10,000 + 1,000Education + \epsilon$  
Set $\beta_0$ to 10,000 and $\beta_1$ to 1,000.

```{r paramters}
beta_0 <-10000 # set intercept
beta_1 <- 1000 # set slope
n <- 100 # set sample size
N <- 100 # set number of simulations
```

#### Simulate 100 draws from the Population Regression Function

$Salary_i = \beta_0 + \beta_1 Education_i + \epsilon_i$

```{r monte-carlo}
results_100 <- data.frame(beta_0 = numeric(), # create a data frame to store the results
                 beta_1 = numeric())
for (i in 1:N){
  x <- sample(0:16, size = n, replace = T) # choose values for x
  e <- rnorm(n, 0, 10000) # generate random errors
  y <- beta_0 + beta_1 * x + e # generate values of y based on the parameter values
  lm(y ~ x) %>% # estimate the model
    broom::tidy() %>% # retrieve the coefficients
    pull(estimate) -> # save the estimates to the data frame
    results_100[i,] 
}
```

### a 

```{r 2a}
results_100 %>% 
  ggplot(aes(x = beta_1)) + 
  geom_histogram(binwidth = 50)

results_100 %>% 
  ggplot(aes(x = beta_1)) +
  geom_density()

results_100 %>% 
  ggplot(aes(y = beta_1)) +
  geom_boxplot()
```

### b

```{r 2b}
results_100 %>% 
  select(beta_1) %>% 
  summarize(beta_1_hat = mean(beta_1),
            minimum = min(beta_1),
            maximum = max(beta_1))
```

### c {.tabset}

#### Simulation with sample size 1000.

```{r 2c-monte-carlo}
beta_0 <-10000 # set intercept
beta_1 <- 1000 # set slope
n <- 1000 # set sample size
N <- 100 # set number of simulations

results_1000 <- data.frame(beta_0 = numeric(), # create a data frame to store the results
                 beta_1 = numeric())
for (i in 1:N){
  x <- sample(0:16, size = n, replace = T) # choose values for x
  e <- rnorm(n, 0, 10000) # generate random errors
  y <- beta_0 + beta_1 * x + e # generate values of y based on the parameter values
  lm(y ~ x) %>% # estimate the model
    broom::tidy() %>% # retrieve the coefficients
    pull(estimate) -> # save the estimates to the data frame
    results_1000[i,] 
}
```

#### Comparisons to simulation above

```{r 2comp}
results_1000 %>% 
  select(beta_1) %>% 
  summarize(beta_1_hat = mean(beta_1),
            minimum = min(beta_1),
            maximum = max(beta_1))

results_100 %>% 
  bind_rows(results_1000, .id = "simulation")  %>%  
  ggplot(aes(x = simulation, y = beta_1)) + 
  geom_boxplot()
```

### d {.tabset}

#### Simulation with sample size 20.

```{r d-monte-carlo}
beta_0 <-10000 # set intercept
beta_1 <- 1000 # set slope
n <- 20 # set sample size
N <- 100 # set number of simulations

results_20 <- data.frame(beta_0 = numeric(), # create a data frame to store the results
                 beta_1 = numeric())
for (i in 1:N){
  x <- sample(0:16, size = n, replace = T) # choose values for x
  e <- rnorm(n, 0, 10000) # generate random errors
  y <- beta_0 + beta_1 * x + e # generate values of y based on the parameter values
  lm(y ~ x) %>% # estimate the model
    broom::tidy() %>% # retrieve the coefficients
    pull(estimate) -> # save the estimates to the data frame
    results_20[i,] 
}
```

#### Comparisons to simulation above

```{r 2d}
results_20 %>% 
  select(beta_1) %>% 
  summarize(beta_1_hat = mean(beta_1),
            minimum = min(beta_1),
            maximum = max(beta_1))

results_100 %>% 
  bind_rows(results_20, .id = "simulation")  %>%  
  ggplot(aes(x = simulation, y = beta_1)) + 
  geom_boxplot()
```

### e {.tabset}

#### Simulation with sd = 500.

```{r e-monte-carlo}
beta_0 <-10000 # set intercept
beta_1 <- 1000 # set slope
n <- 100 # set sample size
N <- 100 # set number of simulations

results_sd_500 <- data.frame(beta_0 = numeric(), # create a data frame to store the results
                 beta_1 = numeric())
for (i in 1:N){
  x <- sample(0:16, size = n, replace = T) # choose values for x
  e <- rnorm(n, 0, 500) # generate random errors
  y <- beta_0 + beta_1 * x + e # generate values of y based on the parameter values
  lm(y ~ x) %>% # estimate the model
    broom::tidy() %>% # retrieve the coefficients
    pull(estimate) -> # save the estimates to the data frame
    results_sd_500[i,] 
}
```

#### Comparisons to original simulation 

```{r 2e}
results_sd_500 %>% 
  select(beta_1) %>% 
  summarize(beta_1_hat = mean(beta_1),
            minimum = min(beta_1),
            maximum = max(beta_1))

results_100 %>% 
  bind_rows(results_sd_500, .id = "simulation")  %>%  
  ggplot(aes(x = simulation, y = beta_1)) + 
  geom_boxplot()
```

### f {.tabset}

#### Simulation with sd = 50,000.

```{r f-monte-carlo}
beta_0 <-10000 # set intercept
beta_1 <- 1000 # set slope
n <- 100 # set sample size
N <- 100 # set number of simulations

results_sd_50000 <- data.frame(beta_0 = numeric(), # create a data frame to store the results
                 beta_1 = numeric())
for (i in 1:N){
  x <- sample(0:16, size = n, replace = T) # choose values for x
  e <- rnorm(n, 0, 50000) # generate random errors
  y <- beta_0 + beta_1 * x + e # generate values of y based on the parameter values
  lm(y ~ x) %>% # estimate the model
    broom::tidy() %>% # retrieve the coefficients
    pull(estimate) -> # save the estimates to the data frame
    results_sd_50000[i,] 
}
```

#### Comparisons to original simulation 

```{r 2f}
results_sd_50000 %>% 
  select(beta_1) %>% 
  summarize(beta_1_hat = mean(beta_1),
            minimum = min(beta_1),
            maximum = max(beta_1))

results_100 %>% 
  bind_rows(results_sd_50000, .id = "simulation")  %>%  
  ggplot(aes(x = simulation, y = beta_1)) + 
  geom_boxplot()
```

### g 

```{r 1g}
N <- 500
results_100_500 <- data.frame(beta_0 = numeric(), # create a data frame to store the results
                 beta_1 = numeric())
for (i in 1:N){
  x <- sample(0:16, size = n, replace = T) # choose values for x
  e <- rnorm(n, 0, 10000) # generate random errors
  y <- beta_0 + beta_1 * x + e # generate values of y based on the parameter values
  ols <- lm(y ~ x)
  results_100_500[i,] <-coef(ols)
}

results_100_500 %>% 
  select(beta_1) %>% 
  summarize(beta_1_hat = mean(beta_1),
            minimum = min(beta_1),
            maximum = max(beta_1))

results_100_500 %>% 
  ggplot(aes(x = beta_1)) +
  geom_density()
```

### All Densities

```{r all-dens}
results_100 %>% 
  bind_rows(list(results_1000, results_20,
            results_sd_500,results_sd_50000,
            results_100_500), .id = "simulation") %>% 
  ggplot(aes(x = beta_1)) +
  geom_density() + 
  facet_wrap(~ simulation, scales = "free_y")
  


```



## 3 {.tabset}

```{r 3data, message=FALSE}
height_wage <- 
  here::here("data", "heightwage_british_males.csv") %>% 
  read_csv()
```

### a

```{r 3a}
height_wage %>% 
  lm(gwage33 ~ height33, .) %>% 
  jtools::summ()
```

### b

```{r 3b}
height_wage %>% 
  ggplot(aes(x = height33, y = gwage33)) + 
  geom_rect(aes(ymin = 400, ymax = 2600, xmin = 65, xmax = 75), fill = "grey") +
  geom_rect(aes(ymin = 0, ymax = 100, xmin = 20, xmax = 36), fill = "grey") +
  geom_point() 
```

### c

```{r 3c}
height_wage %>% 
  filter(gwage33 <= 400, height33 > 40) %>%  
  ggplot(aes(y = gwage33, x = height33)) +
  geom_jitter(alpha = .3)

```

### d

```{r 3d}
jtools::export_summs(
  height_wage %>% 
  lm(gwage33 ~ height33, .), 
  height_wage %>% 
  filter(gwage33 <= 400, height33 > 40) %>%
  lm(gwage33 ~ height33, .),
  model.names = c("Part (a)", "Part (b)")
)
```


### e 

```{r}
jtools::export_summs(
  height_wage %>%
    lm(gwage33 ~ height33, .),
  height_wage %>%
    slice(1:800) %>%
    lm(gwage33 ~ height33, .)
)
```



## 4 {.tabset}

```{r 4data, message=FALSE}
divorce <- 
here::here("data", "DivorceMen.csv") %>% 
  read_csv() %>% 
  janitor::clean_names() %>% 
  bind_rows(
    here::here("data", "DivorceWomen.csv") %>% 
  read_csv() %>% 
  janitor::clean_names()
    )
```

### a

```{r 4a, message=FALSE}
divorce %>% 
  ggplot(aes(x = gender_hours, y = divorce_rate, color = as.factor(male))) + 
  geom_point() +
  xlab("Yearly labor hours for gender") +
  ylab("Divorce rate per thousand") +
  scale_color_discrete(labels = c("Women", "Men")) +
  facet_grid(~ as.factor(male), scales = "free") + 
  theme(legend.title = element_blank(), 
        strip.text.x = element_blank())
```

### b

```{r 4b}
jtools::export_summs(
  divorce %>% 
  filter(male == 1) %>% 
  lm(divorce_rate ~ gender_hours, .),
  divorce %>% 
  filter(male == 0) %>% 
  lm(divorce_rate ~ gender_hours, .),
  model.names = c("Male Model", "Female Model"),
  digits = 5
  )
```

### c

```{r 4c}
divorce %>% 
  filter(male == 1) %>% 
  lm(divorce_rate ~ gender_hours, .) %>% 
  broom::augment() %>% 
  mutate(country = divorce$country[divorce$male==1]) %>% 
  filter(country == "Germany") %>% 
  select(country, .fitted, .resid)
```

### d

```{r 4d}
divorce %>% 
  filter(male == 0) %>% 
  lm(divorce_rate ~ gender_hours, .) %>% 
  broom::augment() %>% 
  mutate(country = divorce$country[divorce$male==0]) %>% 
  filter(country == "Spain") %>% 
  select(country, .fitted, .resid)
```

## 5 {.tabset}

```{r 5data, message=FALSE}
divorce <- 
here::here("data", "DivorceMen.csv") %>% 
  read_csv() %>% 
  janitor::clean_names() %>% 
  bind_rows(
    here::here("data", "DivorceWomen.csv") %>% 
  read_csv() %>% 
  janitor::clean_names()
    )
```

### a

```{r 5a}
divorce %>% 
  ggplot(aes(x = tax_rate, y = gender_hours, color = as.factor(male))) + 
  geom_point() +
  ylab("Yearly labor hours for gender") +
  xlab("Average effective tax rate") +
  scale_color_discrete(labels = c("Women", "Men")) +
  facet_wrap(~ as.factor(male)) + 
  theme(legend.title = element_blank(), 
        strip.text.x = element_blank())
```

### b

```{r 5b}
jtools::export_summs(
  divorce %>% 
    filter(male == 0) %>% 
    lm(gender_hours ~ tax_rate, .),
  divorce %>% 
    filter(male == 1) %>% 
    lm(gender_hours ~ tax_rate, .),
  model.names = c("Women", "Men")
)

```

### c

```{r 5c}
divorce %>% 
  filter(male == 1) %>% 
  lm(gender_hours ~ tax_rate, .) %>% 
  broom::augment() %>% 
  mutate(country = divorce$country[divorce$male==1]) %>% 
  filter(country == "United States") %>% 
  select(country, .fitted, .resid)
```

### d

```{r 5d}
divorce %>% 
  filter(male == 0) %>% 
  lm(gender_hours ~ tax_rate, .) %>% 
  broom::augment() %>% 
  mutate(country = divorce$country[divorce$male==1]) %>% 
  filter(country == "Italy") %>% 
  select(country, .fitted, .resid)
```

